Imputation of Incomplete Multilevel Data with \pkg{mice}

### 2022-04-05 Manuscript/Imputation_of_Incomplete_Multilevel_Data.Rmd
Impute impact data

Analyze the imputations:
```{r pop-ignored-fit, echo=TRUE, message=FALSE, warning=FALSE}
fit_ignored <- imp_ignored %>% 
  with(lme4::lmer(popular ~ 1 + sex + texp + sex:texp + (1 + sex | school))) 
fit_ignored %>% 
  pool() %>% 
  broom.mixed::tidy()
testEstimates(as.mitml.result(fit_ignored), var.comp = TRUE)
```

@@

```{r impact, echo=FALSE, fig.width=6, fig.height=8}
# load data
data("impact", package = "metamisc")
impact_NA <- readRDS("../Data/impact_NA.RDS") #%>% select(!c(hypox, hypots, tsah, edh))

# # descriptive statistics
# by(impact, impact$name, summary) 
# psych::describe(impact)[,c(2:5,8:9)]
# # missingness
plot_md_pat(impact_NA) 
# TODO: rotate variable labels
# TODO: think about nr of clusters: if there's too many, faceting is no option -> some distribution across the clusters like shepley plot
# TODO: make missingness % gradient for clusters (ie the number of clusters for which this variable is missing).
```

```{r impact_md, echo=FALSE}
# md pattern in first cluster
plot_md_pat(impact_NA[impact_NA$name == "TINT", ])
plot_md_pat(impact_NA[impact_NA$name == "APOE", ])
```

@@

```{r impact_cca, echo=FALSE}
# mod_cca <- glmer(mort ~ type + age + motor_score + pupil + ct + (1 | name), family = "binomial", data = impact_NA) 
mod_cca_dummy <- glm(mort ~ type + age + motor_score + pupil + ct + as.factor(name), family = "binomial", data = impact_NA) 
mod_cca_dummy %>% broom.mixed::tidy()

@@

If we run the multilevel model on the incomplete date, we get a warning about unindentifyability. This means that with CCA as missing data method, we cannot trust the estimates. To still obtain some estimates to compare later, we fit a simplified analysis model: using the cluster variable as indicator in the model.

```{r impact_pred}
pred <- quickpred(impact_NA)
pred[pred == 1] <- 2
pred[, "name"] <- -2
pred["mort", ] <- 2
pred[, "mort"] <- 2
diag(pred) <- 0
pred[c("name", "type", "age", "motor_score", "mort"), ] <- 0
pred

@@

```{r imp_impact}
# imp_impact <- impact_NA %>%
#   mutate(name = as.integer(name), motor_score = as.numeric(motor_score)) %>%
#   mice::mice(., m = 2, maxit = 1, method = meth, predictorMatrix = pred)
# look at jomo for categorical variables?
# semi-cont with jomo is not ideal (schafer, '97) because you need 2-step approach
# pmm is better (more efficient) because it will still look for donors (maybe outside of cluster) based on predictive distance, even for very small clusters
# make assumptions of these methods explicit!
```

>>>>

TODO: remove the broom.mixed output, use mitml only

Analyze the imputations:
```{r pop-ignored-fit, echo=TRUE, message=FALSE, warning=FALSE}
fit_ignored <- imp_ignored %>% 
  with(lme4::lmer(popular ~ 1 + sex + texp + sex:texp + (1 + sex | school))) 
testEstimates(as.mitml.result(fit_ignored), var.comp = TRUE)
```
```{r eval=FALSE, include=FALSE}
# fit_ignored %>% 
#   pool() %>% 
#   broom.mixed::tidy()
```

@@

```{r impact, echo=FALSE, fig.width=6, fig.height=8}
# load data
data("impact", package = "metamisc")
impact_NA <- readRDS("../Data/impact_NA.RDS") %>% 
  mutate(name = as.integer(name),
         motor_score = as.numeric(motor_score))#%>% select(!c(hypox, hypots, tsah, edh))

# # descriptive statistics
# by(impact, impact$name, summary) 
# psych::describe(impact)[,c(2:5,8:9)]
# # missingness
plot_pattern(impact_NA, rotate = TRUE) 
# TODO: think about nr of clusters: if there's too many, faceting is no option -> some distribution across the clusters like shepley plot
# TODO: make missingness % gradient for clusters (ie the number of clusters for which this variable is missing).
```

```{r impact_md, echo=FALSE}
# # md pattern in first cluster
# plot_pattern(impact_NA[impact_NA$name == "TINT", ])
# plot_pattern(impact_NA[impact_NA$name == "APOE", ])
```

@@

```{r impact_cca, echo=FALSE}
mod_cca <- glmer(mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 | name), family = "binomial", data = impact_NA) 
mod_cca %>% broom.mixed::tidy()
# mod_cca_dummy <- glm(mort ~ type + age + motor_score + pupil + ct + name, family = "binomial", data = impact_NA) 
# mod_cca_dummy %>% broom.mixed::tidy()

@@

<!-- If we run the multilevel model on the incomplete date, we get a warning about unindentifyability. This means that with CCA as missing data method, we cannot trust the estimates. To still obtain some estimates to compare later, we fit a simplified analysis model: using the cluster variable as indicator in the model. -->

We should impute the variables `ct` and `pupil` and any auxiliary variables we might want to use to impute these incomplete analysis model variables. We can evaluate which variables may be useful auxiliaries by plotting the pairwise complete correlations:

```{r impact_corr}
plot_corr(impact_NA, label = T)
```

This shows us that `hypox` and `hypot` would not be useful auxiliary variables for imputing `ct`. Depending on the minimum required correlation, `tsah` could be useful, while `edh` has the strongest correlation with `ct` out of all the variables in the data and should definitely be included in the imputation model. For the imputation of `pupil`, none of the potential auxiliary variables has a very strong relation, but `hypots` could be used. We conclude that we can exclude `hypox` from the data since this is neither an analysis model variable, nor an auxiliary variable for imputation.

```{r impact-pred}
impact_NA <- select(impact_NA, -c(hypox, hypots))
pred <- quickpred(impact_NA)
plot_pred(pred)

pred[pred == 1] <- 2
pred["mort", ] <- 2
pred[, "mort"] <- 2
pred[c("name", "type", "age", "motor_score", "mort"), ] <- 0
pred[, "name"] <- -2
diag(pred) <- 0
plot_pred(pred)

@@

```{r imp_impact}
imp <- mice(impact_NA, method = meth, predictorMatrix = pred, m = 2, maxit = 1)
# imp_impact <- impact_NA %>%
#   mutate(name = as.integer(name), motor_score = as.numeric(motor_score)) %>%
#   mice::mice(., m = 2, maxit = 1, method = meth, predictorMatrix = pred)
# look at jomo for categorical variables?
# semi-cont with jomo is not ideal (schafer, '97) because you need 2-step approach
# pmm is better (more efficient) because it will still look for donors (maybe outside of cluster) based on predictive distance, even for very small clusters
# make assumptions of these methods explicit!
```
```{r impact-fit}
fit <- imp_impact_pmm %>% 
  with(glmer(mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 | name), family = "binomial")) 
tidy(pool(fit))
as.mitml.result(fit)
# testEstimates(as.mitml.result(fit))
```

Compare estimates to complete data

```{r}
fit <- glmer(mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 | name), family = "binomial", data = impact) 
tidy(fit)
```

### 2022-04-05 Manuscript/Imputation_of_Incomplete_Multilevel_Data.tex
Impute impact data

\begin{CodeChunk}
\begin{CodeInput}
R> set.seed(2022)
R> library(mice)
R> library(ggmice)
R> library(ggplot2)
R> library(dplyr)
R> library(lme4)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> est_cca <- lme4::lmer(popular ~ 1 + sex + texp + sex:texp + (1 + sex | school), popmis) 
\end{CodeInput}
\end{CodeChunk}

@@

Create a predictor matrix and remove \texttt{school} as predictor for
\texttt{popular}:

\begin{CodeChunk}
\begin{CodeInput}
R> # predictor matrix
R> pred <- quickpred(popmis)
R> plot_pred(pred)
\end{CodeInput}

\begin{center}\includegraphics{Imputation_of_Incomplete_Multilevel_Data_files/figure-latex/pop-ignored-1} \end{center}

\begin{CodeInput}
R> # impute the data, ignoring the cluster structure
R> imp_ignored <- mice(popmis, maxit = 1, pred = pred, print = FALSE)
\end{CodeInput}
\end{CodeChunk}

>>>>

\begin{CodeChunk}
\begin{CodeInput}
R> set.seed(2022)
R> library(mice)
R> library(ggmice)
R> library(ggplot2)
R> library(dplyr)
R> library(lme4)
R> library(mitml)
\end{CodeInput}
\end{CodeChunk}

@@

@@

Create a methods vector and predictor matrix for \texttt{popular}, and
make sure \texttt{school} is not included as predictor:

\begin{CodeChunk}
\begin{CodeInput}
R> meth <- make.method(popmis) # methods vector
R> pred <- quickpred(popmis)   # predictor matrix
R> plot_pred(pred)
\end{CodeInput}

\begin{center}\includegraphics{Imputation_of_Incomplete_Multilevel_Data_files/figure-latex/pop-ignored-pred-1} \end{center}

Impute the data, ignoring the cluster structure:

\begin{CodeChunk}
\begin{CodeInput}
R> imp_ignored <- mice(popmis, maxit = 1, pred = pred, print = FALSE)
\end{CodeInput}
\end{CodeChunk}

TODO: remove the broom.mixed output, use mitml only

Analyze the imputations:

\begin{CodeChunk}
\begin{CodeInput}
R> fit_ignored <- imp_ignored %>% 
+   with(lme4::lmer(popular ~ 1 + sex + texp + sex:texp + (1 + sex | school))) 
R> testEstimates(as.mitml.result(fit_ignored), var.comp = TRUE)
\end{CodeInput}
\begin{CodeOutput}

Call:

testEstimates(model = as.mitml.result(fit_ignored), var.comp = TRUE)

Final parameter estimates and inferences obtained from 5 imputed data sets.

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     3.354     0.245    13.704     6.831     0.000     3.260     0.813 
sex             1.247     0.332     3.759     6.154     0.009     4.160     0.849 
texp            0.110     0.017     6.320     6.084     0.001     4.286     0.852 
sex:texp       -0.027     0.022    -1.225     5.914     0.267     4.631     0.862

                            Estimate 
Intercept~~Intercept|school    0.177 
sex~~sex|school                0.232 
Intercept~~sex|school         -0.043 
Residual~~Residual             0.670 
ICC|school                     0.209

Unadjusted hypothesis test as appropriate in larger samples.
\end{CodeOutput}
\end{CodeChunk}
